---
title: "Clean_Data_to_NLP"
author: "Leena Sonpethkar"
date: "April 8, 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## SwiftKey Capstone Project 
### Synopsis

The objectives for this project is to: 

1. Demonstrate given data has been downloaded and loaded successfully for project

2. Basic report of summary statistics about the data sets. 

3. Report any interesting findings that are found so far.

In my project, I will be performing the following steps: 

1. Download the data and load the data 

2. Pull random samples of the data i.e. sampling which will reduce size significantly and also help in faster processing. 

3. Some initial cleaning of the data. 

4. Pull some basic statistics 

5. Create some simple bar plots and word clouds for 1-4 word sequences 

6. Summary my findings and talk about what I will likely do going forward thru the rest of the class.

```{r, echo=FALSE, message=FALSE}
library(dplyr)
library(R.utils)
library(stringi)
library(tm)
library(RWeka)
library(ggplot2)
library(wordcloud2)
```

```{r, echo=TRUE, message=TRUE}
zipUrl <- "https://d396qusza40orc.cloudfront.net/dsscapstone/dataset/Coursera-SwiftKey.zip"
zipFile <- paste("Data", "Coursera-SwiftKey.zip", sep = "/")

# Download data/zip file 
if (!file.exists(zipFile)) {
  download.file(zipUrl, zipFile, mode = "wb")
}
#Set Path file names
dataPath <- "Data/final/en_US"
infoBL <- paste(dataPath, "en_US.blogs.txt", sep = "/")
infoNW <- paste(dataPath, "en_US.news.txt", sep = "/")
infoTW <- paste(dataPath, "en_US.twitter.txt", sep = "/")

# Unzip data file 
if (!file.exists(file.path(infoBL))) {
  unzip(zipFile, exdir = "/Data")
}
```
### Basic report of summary statistics about the data sets. 
To understand the depth of the dataset, let’s run some basic statistics on the text in the 3 files (blogs, news & twitter). We will calculate the total number of lines, characters and words in the data as well as minimum, average (mean) and maximum word counts for any the 3 data files.

```{r, echo=TRUE}
#Open connection and Read Data
conBL <- file(file.path(paste(getwd(), infoBL, sep = "/")), "r") 
conNW <- file(file.path(paste(getwd(), infoNW, sep = "/")), "r") 
conTW <- file(file.path(paste(getwd(), infoTW, sep = "/")), "r") 

rdBL <- readLines(conBL, warn=FALSE, encoding="UTF-8", skipNul = TRUE)
rdNW <- readLines(conNW, warn=FALSE, encoding="UTF-8", skipNul = TRUE)
rdTW <- readLines(conTW, warn=FALSE, encoding="UTF-8", skipNul = TRUE) 

#size
infoBL[2] <- file.info(paste(getwd(), infoBL, sep = "/"))$size/1048576
infoNW[2] <- file.info(paste(getwd(), infoNW, sep = "/"))$size/1048576
infoTW[2] <- file.info(paste(getwd(), infoTW, sep = "/"))$size/1048576

#No. of lines
infoBL[3] <- length(rdBL)
infoNW[3] <- length(rdNW)
infoTW[3] <- length(rdTW)

#nu. of chars
infoBL[4] <- sum(stri_count_words(rdBL))
infoNW[4] <- sum(stri_count_words(rdNW))
infoTW[4] <- sum(stri_count_words(rdTW))

#max length
infoBL[5] <- max(nchar(rdBL))
infoNW[5] <- max(nchar(rdNW))
infoTW[5] <- max(nchar(rdTW))


df1 <- data.frame(matrix(c(infoBL,infoNW, infoTW),nrow = 3, ncol = 5, byrow = TRUE))
colnames(df1) <- c("file name", "size(MB)", "num.of.lines", "num.of.chars", "longest.line")
df1
```
###Sampling the data
Random sample of 3% of data from each file will be used for initial analysis.
This is to allow us to provide a faster look at the data to see what we see and decide on the next steps before analyzing the entire data files.

A Corpus is created based on the 3 samples.
```{r, echo=TRUE}
set.seed(190421)

dtBL_sample <- sample(rdBL, length(rdBL)*.0)
dtNW_sample <- sample(rdNW, length(rdNW)*.0)
dtTW_sample <- sample(rdTW, length(rdTW)*.75)

combined_sample <- c(dtBL_sample, dtNW_sample, dtTW_sample)
combined_sample <- iconv(combined_sample, "UTF-8","ASCII", sub="")

length(combined_sample)

#Clear memory by removing unwanted dataframes
#rm(dtBL_sample, dtNW_sample, dtTW_sample, rdBL, rdNW, rdTW)
rm(dtBL_sample, dtNW_sample, rdBL, rdNW, rdTW)
```

```{r, echo=TRUE}
#combined_sample <- iconv(rdTW, "UTF-8","ASCII", sub="")


corpus <- VCorpus(VectorSource(combined_sample))
corpus <- tm_map(corpus, tolower)  #1
corpus <- tm_map(corpus, removePunctuation) #2
corpus <- tm_map(corpus, removeNumbers) #3
corpus <- tm_map(corpus, removeWords, stopwords("english")) #4
corpus <- tm_map(corpus, PlainTextDocument) #5
corpus <- tm_map(corpus, stripWhitespace) #6

#file.path(paste(getwd(), infoBL, sep = "/"))

#tidy_comb_sample = paste(dataPath, "tidy_comb_sample_en.rds", sep = "/")

#writeCorpus(corpus, path = dataPath, filenames = paste("tidy_comb_sample_en", ".txt", sep = ""))

#saveRDS(corpus, file=tidy_comb_sample)
```

```{r, echo=TRUE}


trigram <- function(x) NGramTokenizer(x, Weka_control(min = 3, max = 3))
                                      #, delimiters = "\\r\\n\\t.,;:\"()?!")
trigram_tdm <- TermDocumentMatrix(corpus, control = list(tokenize = trigram))
#trigram_freqTerm <- findFreqTerms(trigram_tdm,lowfreq=50)
#trigram_freq <- rowSums(as.matrix(trigram_tdm[trigram_freqTerm,]))
#trigram_ord <- order(trigram_freq, decreasing = TRUE)
#trigram_freq <- data.frame(word=names(trigram_freq[trigram_ord]), frequency=trigram_freq[trigram_ord])

triGramDf <- data.frame(table(trigram_tdm)) 

sanitizeGramDf <- function(df) { 
newDf <- data.frame(Term = as.character(df[, 1]), Count = df[, 2]) 
newDf 
} 
  
triGramDf <- sanitizeGramDf(triGramDf) 
  
sortGramDf <- function(df) { 
 df[order(df$Count, decreasing = TRUE), ] 
 } 
  
triGramDf <- sortGramDf(triGramDf) 
 
```


```{r echo=TRUE}
## Other Sample code

gramTokenizer <- function(n) { 
NGramTokenizer(ovid, Weka_control(min = n, max = n, delimiters = " \\r\\n\\t.,;:\"()?!")) 
} 

  
oneGram <- gramTokenizer(1) 
biGram <- gramTokenizer(2) 
triGram <- gramTokenizer(3) 
  
oneGramDf <- data.frame(table(oneGram)) 
biGramDf <- data.frame(table(biGram)) 
triGramDf <- data.frame(table(triGram)) 
  
sanitizeGramDf <- function(df) { 
newDf <- data.frame(Term = as.character(df[, 1]), Count = df[, 2]) 
newDf 
} 
  
 oneGramDf <- sanitizeGramDf(oneGramDf) 
 biGramDf <- sanitizeGramDf(biGramDf) 
 triGramDf <- sanitizeGramDf(triGramDf) 
  
 sortGramDf <- function(df) { 
 df[order(df$Count, decreasing = TRUE), ] 
 } 
  
 oneGramDf <- sortGramDf(oneGramDf) 
 biGramDf <- sortGramDf(biGramDf) 
 triGramDf <- sortGramDf(triGramDf) 
  
 reductionRows <- c(1: 30) 
 oneGramDfReduced <- oneGramDf[reductionRows, ] 
 biGramDfReduced <- biGramDf[reductionRows, ] 
 triGramDfReduced <- triGramDf[reductionRows, ] 

```
